---
title: 'Simulation results: summary tables, bias and precision'
author: "Ann Von Holle"
date: "October 20, 2015"
output: html_document
csl: american-journal-of-epidemiology.csl
bibliography: bib1.bib
---

This document is based on ~/Dropbox/unc.grad.school/my-papers/ms-201505/programs/read-mixsim.growth-8-sim-subset.Rmd 

Will provide a mock-up of figures meant for poster (before putting in .Rnw file)

Based on analyses for the manuscript currently titled, "The utility of standardized or crude weight measures in modeling of postnatal growth trajectories: Are there differences?". The aim of the paper is to compare growth trajectory models with three different outcomes: weight, weight Z-score and weight percentile. This topic has been discussed in the literature [@berkey_adiposity_2007,@cole_what_2005], but not for the infancy time period.


---

### Models used for evaluation across the three outcomes:

#### Model 1: Linear regression with first order term for time.

$y_{ij} = \beta_0 + \beta_1 \cdot t + \beta_2 \cdot \textrm{group} + \beta_3 \cdot t \cdot \textrm{group} + e_{ij}$

#### Model 2: Linear regression with second order term for time.

$y_{ij} = \beta_0 + \beta_1 \cdot t + \beta_2 \cdot \textrm{group} + \beta_3 \cdot t \cdot \textrm{group} + \beta_4 \cdot t^2 + e_{ij}$


#### Model 3: ANOVA model (month 6 versus month 0).

$y_{ij} = \beta_0 + \beta_1 \cdot month.6 + \beta_2 \cdot \textrm{group} + \beta_3 \cdot month.6 \cdot \textrm{group} + e_{ij}$

$month.6$ is a binary variable with 1=month 6 and 0=month 0.



```{r global_options, include=F}
# see http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, 
                      results='hide')
```

```{r}
# package info
library(knitr)
library(reshape2)
library(ggplot2)
library(car) 
library(plyr)
library(tables)
library(data.table)
library(htmlTable)
require(ztable)
require(Rmisc)

library(gridExtra)
library(grid)
```

```{r read1}
#setwd("C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201505/programs")
setwd("~/")
setwd("../")
setwd("./Dropbox/unc.grad.school/my-papers/ms-201505/programs")

# Read in data and get power and nominal alpha
# ................................................

load("mixsim.growth-8-sim-subset.Rda") # this file has the vals.1 object from mixsim.growth-8-sim-subset.R. Run on killdevil.

head(vals.1) # note that all data is in tails -- females only
dim(vals.1)
class(vals.1) # matrix
colnames(vals.1)
head(vals.1)


# different way to split vals.b and get mean for columns 1 to 18 for vals.b
# ...................................................................
vals.b.dt = data.table(vals.1)
names(vals.b.dt)
dim(vals.b.dt)

table(vals.b.dt$study)
```


```{r handle1}
# Get proportion significant from samples
# ............................................
colnames(vals.b.dt)

dat.1 = vals.b.dt[,lapply(.SD, function(x) {formatC( round(mean(ifelse(x<0.05, 1, 0), na.rm=T),3), format="f", digits=3)}),
                  by = c("sigval","study", "gend", "basevals"),
                  .SDcols=c(1:18,55,56)]

#head(dat.1)
#levels(factor(dat.1$e))
dim(dat.1)
```

```{r handle2}
# Get median of the coefficient estimates for the interaction term
# ...............................................................

names(vals.b.dt)

# see http://www.magesblog.com/2012/01/say-it-in-r-with-by-apply-and-friends.html
dat.2.median = vals.b.dt[,lapply(.SD, function(x) {median(x, na.rm=T)}),
                  by = c("sigval","study", "gend", "basevals"),
                  .SDcols=c(19:36,57,58)]
dat.2.median$measure="median"
dat.2.median

dat.2.iqr = vals.b.dt[,lapply(.SD, function(x) {IQR(x, na.rm=T)}),
                  by = c("sigval","study", "gend", "basevals"),
                  .SDcols=c(19:36,57,58)]
dat.2.iqr$measure="median"
dat.2.cont = rbind(dat.2.median,dat.2.iqr)
head(dat.2.cont)

# Get median of sample regresssion interaction coefficients and sd of sample regression int. coefficients
# ..........................................
dat.2a = vals.b.dt[,lapply(.SD, function(x) {median(x, na.rm=T)}),
                  by = c("sigval","study", "gend", "basevals"),
                  .SDcols=c(19:36,57,58)]
names(dat.2a)
head(dat.2a)
```

```{r handle3}
# More data handling
# .........................................

studies = c("Portugal", "Italy", "Chile") # fix study labels for formatting below.
evals = c("Unequal baseline \n + slope",
          "Equal baseline \n + slope",
          "Equal baseline \n - slope",
          "Unequal baseline \n - slope")

# Combine the means with alpha and power vals
dat.both = cbind(data.frame(dat.1), data.frame(dat.2a[,5:24, with=F]))
levels(factor(dat.both$basevals))

# Change the data from wide to long so I can make a table
# .........................................................
names(dat.both)

dat.long = melt(dat.both[,c(1:44)], 
                id.vars=c("sigval","study", "gend", "basevals"))

names(dat.long)
head(dat.long)
tail(dat.long)

handle.1 = function(df){
within(df, {
  power = ifelse(grepl("null.pval",variable)==T,"Type I error",
                 ifelse(grepl("alt.pval", variable)==T, "Power",
                        ifelse(grepl(".val", variable)==T, "Interaction coefficient: median (IQR)^a", NA)))
  type = ifelse(grepl("zwei.", variable)==T, "Z",
                ifelse(grepl("wt.", variable)==T, "Weight",
                       ifelse(grepl("pct.", variable)==T, "Percentile", NA)))
    alt = ifelse(grepl("alt.", variable)==T, 1, 0)
    val.1 = ifelse(grepl("alt.va", variable)==T, 1, 0)

  model = ifelse(grepl("m3", variable)==T, "3",
                ifelse(grepl("m2", variable)==T, "2", 
                       ifelse(grepl("m1", variable)==T, "1",
                              ifelse(grepl("mg", variable)==T, "Source", NA))))
  #value.2 = round(value, 3)
  #gender.f = factor(gender, labels=c("Male", "Female"))
  gender.f = factor(gend, labels=c("Female")) 
#  population.f = factor(population, labels=c("South Africa", "Malawi", "Portugal", "Italy", "Chile"))
  population.f = factor(study, labels=studies) 
#  population.f = factor(study, labels=c("Portugal")) 
   sigma.f = factor(sigval, labels=c("0.50"))
   e.f = factor(basevals, labels=evals)
})
}

# get levels so I can pick appropriate formatting above.
names(dat.long)
sapply(dat.long[,c(1,2,3,4)], function(x) levels(factor(x)))

dat.long=handle.1(dat.long)

names(dat.long)
nrow(dat.long)

# reorder
dat.long$power.f = factor(dat.long$power, levels=c("Power", 
                                                   "Type I error", 
                                                   "Interaction coefficient: median (IQR)^a")) # re-order values in way they will appear in table
levels(dat.long$power.f)

dim(dat.long)
head(dat.long)
tail(dat.long)
dat.long[dat.long$model=="Source", ]

```

<!---
Prep data for performance of bias and estimates of uncertainty
See http://www4.stat.ncsu.edu/~davidian/st810a/,
Slides titled, "Simulation studies in statistics" by Dr. Davidian
-->

```{r prep1}

mu = 0.5 # the fixed value of mu set in the simulation -- mixsim.growth-8-sim-subset.R
slope=0.5 # value for slope in simulation

colnames(vals.b.dt[,28:33, with=F])
# Note: each column corresponds to an estimate and each row corresponds to a sample

vals.b.dt = within(vals.b.dt, {
  sigma=sigval
  population.f = factor(study, labels=studies) 
  e.f = factor(basevals, labels=evals)
  gender=gend
})

colnames(vals.b.dt)
# pval: 1:18,55,56; vals: 19:36,57,58; se: 37:54,59,60

# get sample mean estimate (of beta5, the interaction term) by groups (population, gender and fixed effect values) over each of the xxx iterations per group

table(vals.b.dt$basevals) # basevals 1 and 2 have positive slope of 0.5 and basevals 3 and 4 have negative slope of -0.5; See baseint matrix in mixsim.growth-8-sim-subset.R

vals.b.dt[,bias.neg:=lapply(.SD, function(x) {mean(x, na.rm=T)-(-0.5)}),
                  by = c("sigma","population.f", "gender", "e.f", "basevals"),
                  .SDcols=c(19:36,57,58)]

vals.b.dt[,bias.pos:=lapply(.SD, function(x) {mean(x, na.rm=T)-(0.5)}),
                  by = c("sigma","population.f", "gender", "e.f", "basevals"),
                  .SDcols=c(19:36,57,58)]

vals.b.dt$value = with(vals.b.dt, {
  ifelse(basevals %in% c(1,2), bias.pos, bias.neg)
})

bias = as.data.frame(vals.b.dt)
bias$type="Bias"

# Example:
# see http://www4.stat.ncsu.edu/~davidian/st810a/mean.R
#  function to generate S data sets of size n from normal
#  distribution with mean mu and variance sigma^2

# bind together and melt to extract out variable name info

bias.l = melt(bias, id.vars=c("sigma", "population.f",
                                       "gender", "e.f", "type", "basevals"))

info.1 = bias.l

info.1 = within(info.1, {
  outcome = ifelse(grepl("zwei.", variable)==T, "Z",
                ifelse(grepl("wt.", variable)==T, "Weight",
                       ifelse(grepl("pct.", variable)==T, "Percentile", NA)))
  model = ifelse(grepl("m3", variable)==T, "3",
                ifelse(grepl("m2", variable)==T, "2", 
                       ifelse(grepl("m1", variable)==T, "1",
                              ifelse(grepl("mg", variable)==T, "Source", NA))))
  alt = ifelse(grepl(".alt", variable)==T, 1,0)
})

info.1 = info.1[info.1$alt==1,] # alt=1 has a scenario with real difference
head(info.1)

```

---

Plots for the poster go here...

```{r powerplot}
dat.long$study.f = factor(dat.long$study,
                          labels=c("Portugal",
                                   "Italy",
                                   "Chile"))
levels(factor(dat.long$type))
dat.long$type.f = factor(dat.long$type, levels=c("Weight", "Z", "Percentile"))
levels(dat.long$type.f)

# Power ------------------------------------------
table(dat.long$power.f)

head(dat.long[dat.long$power.f=="Interaction coefficient: median (IQR)^a",])

dat.power = dat.long[dat.long$power.f=="Power" & dat.long$alt==1,]
head(dat.power)
dat.power = dat.power[complete.cases(dat.power),]
head(dat.power)
dat.power$value = as.numeric(dat.power$value)
head(dat.power)

g.1 <- guide_legend("Outcome")
g.2 <- guide_legend("Sample")

p.1 = ggplot(aes(x=model, y=value, colour=type.f, shape=study.f), 
             data=dat.power) +
  geom_point(size=6) +
  facet_grid(e.f ~.) + 
  theme_bw(base_size = 26) +
  xlab("Model") + ylab("") +
  ggtitle("Power") +
  scale_y_continuous(breaks=pretty(dat.power$value, n = 5)) +
  guides(colour = g.1,
         shape = g.2) +
  theme(legend.position="bottom",
        strip.text.y = element_text(size = 20))

p.1

# Type I error ------------------------------------------------------

dat.t1e = dat.long[dat.long$power.f=="Type I error",]
head(dat.t1e)
dat.t1e = dat.t1e[complete.cases(dat.t1e),]
head(dat.t1e)
dat.t1e$value = as.numeric(dat.t1e$value)

p.2 = ggplot(aes(x=model, y=value, colour=type.f, shape=study.f), data=dat.t1e) +
  geom_point(size=6) +
  facet_grid(e.f ~.) + 
  theme_bw(base_size = 26) +
  geom_hline(yintercept=0.05, lty=3, lwd=2, colour="red") +
  xlab("Model") + ylab("") +
  ggtitle("Type I error") +
  scale_y_continuous(breaks=pretty(dat.t1e$value, n = 4)) +
  guides(colour = g.1,
         shape = g.2) +
  theme(legend.position="bottom",
        strip.text.y = element_text(size = 20))

p.2
  
# BIAS ----------------------------------------

#info.1.sub=info.1[info.1$type=="Bias" & !(info.1$outcome %in% c("Percentile")),]
info.1.sub=info.1[info.1$type=="Bias",]
table(info.1.sub$outcome)

info.1.sub = info.1.sub[complete.cases(info.1.sub),]
head(info.1.sub)
table(info.1.sub$e.f)

# Use dat.long data frame (instead of info.1)
dat.bias = dat.long[dat.long$power.f=="Interaction coefficient: median (IQR)^a" & dat.long$alt==1,]
# NOTE: need to go back and make sure I want the alt=1 and not alt=0
dat.bias = dat.bias[complete.cases(dat.bias),]
names(dat.bias)
dat.bias$outcome = dat.bias$type

dat.bias[,c("value", "population.f", "model", "alt", "e.f", "outcome")]

class(dat.bias$value)
dat.bias$value = as.numeric(dat.bias$value)
table(dat.bias$basevals) # basevals 1,2 are positive slope, 0.5 and basevals 3,4 are negative slope, -0.5. See mixsim.growth-8-sim-subset.R base.int matrix object.
head(dat.bias$value) 
dat.bias$bias = with(dat.bias, {
  ifelse(basevals %in% c(1,2), value-0.05, value+0.05)
})

table(dat.bias$model)

names(dat.bias)

dat.bias[1:100, c("study","basevals","value","e.f","model","type.f","outcome","study.f")]

summary(dat.bias$bias)

# | dat.bias$type=="Percentile"
sub.dat.bias = dat.bias[!( dat.bias$model==3 | dat.bias$model=="Source"),]
table(sub.dat.bias$model)

p.3 = ggplot(aes(x=model, y=bias, colour=type.f, shape=study.f), data=sub.dat.bias) +
  geom_point(size=6) +
  facet_grid(e.f ~., scales="free") + 
  theme_bw(base_size = 26) +
  xlab("Model") + ylab("") +
  geom_hline(yintercept=0, lty=3, lwd=2, colour="red") +
  ggtitle("Bias") +
  scale_y_continuous(breaks=pretty(sub.dat.bias$bias, n = 5)) +
  ylim(-0.1, 0.1)+
  guides(colour = g.1,
         shape = g.2) +
  theme(legend.position="bottom",
        strip.text.y = element_text(size = 20))

p.3

# see http://stackoverflow.com/questions/13649473/add-a-common-legend-for-combined-ggplots
# for info on adding a legend to bottom of plot (and code below that I copied)

grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    grid.arrange(
        do.call(arrangeGrob, lapply(plots, function(x)
            x + theme(legend.position="none"))),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight))
}

grid_arrange_shared_legend(p.1, p.2, p.3)
#multiplot(p.1, p.2, p.3, cols=3)

# see http://www.sthda.com/english/wiki/ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page-r-software-and-data-visualization

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(p.1)

grid.arrange(p.1 + theme(legend.position="none"),
             p.2 + theme(legend.position="none"),
             p.3 + theme(legend.position="none"),
             legend, 
             ncol=3, nrow=2, 
             layout_matrix = rbind(c(1,2,3), c(4,4,4)),
             widths = c(2.7, 2.7, 2.7), 
             heights = c(2.5, 0.2))

setwd("~/")
setwd("../")
getwd()
setwd("./Dropbox/unc.grad.school/presentations/Epi-congress-of-americas-2016/ser-poster-201606")

# export to png TO GO INTO THE POSTER, MADE AT SER-2016-POSTER-AVH.RNW
png("plot-power.png", width=5800, height=4300, res=300)
 
grid.arrange(p.1 + theme(legend.position="none"),
             p.2 + theme(legend.position="none"),
             p.3 + theme(legend.position="none"),
             legend, 
             ncol=3, nrow=2, 
             layout_matrix = rbind(c(1,2,3), c(4,4,4)),
             widths = c(2.7, 2.7, 2.7), 
             heights = c(2.5, 0.2))
dev.off()

```


